---
output:
  word_document: default
  html_document: default
---

&nbsp;

# Lucas de Paula
# Module 4 Assignment 1

&nbsp;

```{r, include = FALSE}

library(tidyverse)
library(mice)
library(skimr)
library(VIM)
```

```{r ReadData}

grades = read.csv("data/class-grades.csv")

```

&nbsp;

The dataset was originally used to investigate how student performance in the course would be predictive of student grades on the “Final” exam. 
The “Prefix” column is a surrogate for enrollment year in the engineering program.  Smaller values imply older (more mature?) students.

&nbsp;

# Task 1
## For which variables is there missing data?

&nbsp;

```{r MissingAnalysis}

skim(grades)

```

&nbsp;

As we can see above, the variables Tutorial, Midterm, TakeHome and Final have missing values.

&nbsp;

# Task 2: 
## Use the VIM package to visualize missingness. Are there any students that are missing multiple pieces of data?

&nbsp;

```{r Task2}

vim_plot = aggr(grades, numbers = TRUE, prop = c(TRUE, FALSE), cex.axis=.7)

```

&nbsp;

Yes, there is one student that has TakeHome and Midterm containing missing values.

&nbsp;

# Task 3: 
## Use row-wise deletion of missing values to create a new data frame. How many rows remain in this data frame?

&nbsp;

```{r RowWise}

rowwise_grades = grades %>% drop_na() 

nrow(rowwise_grades)

```

&nbsp;

There are 89 observations on the newly created data frame.

&nbsp;

# Task 4: 
## Use column-wise deletion of missing values to create a new data frame (from the original dataframe not from the data frame created in Task 3). 
## How many columns remain in this data frame?

&nbsp;

```{r ColWise}

colwise_grades = grades %>% select(-Tutorial, -Midterm, -TakeHome, -Final)

ncol(colwise_grades)


```

&nbsp;

There are only two columns left.

&nbsp;

# Task 5: 
## Which approach (Task 3 or Task 4) seems preferable for this dataset? Briefly discuss your answer.

&nbsp;

I'd say the first approach (row-wise deletion) makes more sense as we are losing much less information during the selection.
Although we lose 10 rows of data, it means minimal loss of information when compared to losing 4 out of 6 columns.

&nbsp;

# Task 6:
## Use the “mice” package to impute the missing data in this dataset. Set “m = 5” and “method=”pmm"" in the imputation. 
## You MUST use a random number seed of 123 for this imputation. 
## What is the mean value for the “Final” variable after imputation?

&nbsp;

```{r ImputeMissing}

set.seed(123) 

imp_grades = mice(grades, m=5, method='pmm', printFlag=FALSE)

summary(imp_grades)

grades_complete = complete(imp_grades) 

summary(grades_complete)

```

&nbsp;

The mean value for the Final variable after imputation is 68.33.

&nbsp;

